---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# Overview

This repository is connected with the final assignment for __Building R Packages__ module of the __[Mastering Software Development in R](https://www.coursera.org/specializations/r)__ specialization (coursera). The goal is to apply the basic concepts learnt in the module for creating, writing, documenting, and testing an R package. 

The development of the package is done introducing the following software engineering practices:

* version control using Github/ Git
* Test Driven Development with unit tests and integration tests
* Continuous Integration (CI) using [Travis CI](https://travis-ci.org)

## What needs to be done

* add the FARS R code provided during the course
  * refactor R code in order to remove the __Non-Standard Evaluation Issue__ [dplyr NSE vignette](https://cran.r-project.org/web/packages/dplyr/vignettes/nse.html), [NSE explained](http://adv-r.had.co.nz/Computing-on-the-language.html).
* document the code, creating related help functions with `roxygen2`
* document the package, using a README file and a vignette to include in your package with `knitr` and __R Markdown__
* add some unit tests / integration tests using `testthat`
* add the package on GitHub
* set up the repository so that the package can be checked and built on Travis with no errors, warnings or notes.

__IMPORTANT!!__ The "maps" package has been added in the _Depends_ section in the __DESCRIPTION__ file because it sets some environment variables when its loaded & attached, which doesn't happen if maps::map() is used throwing a  "stateMapEnv not found" error.

# `farsr` package ![Travis Badge](https://travis-ci.org/pparacch/fars.svg?branch=master)

## Overview

The goal of `fars` is to help you performing basic operation on [Fatality Analysis Reporting System](https://www.nhtsa.gov/research-data/fatality-analysis-reporting-system-fars) (FARS) data, a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes. Specifically it allows you to perform the following operations:

* load the FARS datafile by filename or by years,
* summarize the number of fatal injuries suffered in motor vehicle traffic crashes by month and year,
* visualize the fatal injuries suffered in motor vehicle traffic crashes in a specific state (USA) for a specific year.

## Installation

```{r eval=FALSE}
#Using devtools and Github
devtools::install_github("pparacch/fars")
```

## Getting Started

```{r eval=FALSE}
#load and attach the package
library(farsr)
```

Use the following table to have an overview of the provided functions:

| Function  |  |
| :--------- | :-------------- |
| `fars_read()`  | Read a FARS datafile |
| `fars_read_years()` | Read FARS datafile(s) for provided year(s) |
| `fars_summarize_years()` | Summarizes injuries by month for provided year(s) |
|`fars_map_state()` | Visualize map with injuries for provided state (USA) and year |

More info can be found in the package vignette `vignette('introduction', package = 'farsr')` and help of provided function e.g. `?farsr::fars_read_years`.






